---
title: "Team_Alpha_Task 2"
author: "Team Alpha"
date: "23/09/2021"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

## Introduction

What is the Benefits data set about? The benefits data set contains a
cross section from 1972 of unemployed blue-collar workers.

Data and the description can be found
[here](<http://www-eio.upc.edu/~pau/cms/rdata/datasets.html>).

Several of the variable headings were not clearly defined and some of
the terminologies were not familiar. Below are the definition of some of
the terms used.

Unemployment Insurance program -

Defining some of the concepts of Unemployment Insurance payments
(benefits) - <https://www.dol.gov/general/topic/unemployment-insurance>;
<https://oui.doleta.gov/unemploy/uifactsheet.asp> state unemployment
rate, - state maximum benefit level, -
<https://www.healthinsurance.org/glossary/maximum-dollar-limit/>

Job Lost, - blue collar workers, -
<https://www.investopedia.com/terms/b/bluecollar.asp> smsa,

-   job displacement, -
    <https://www.mightyrecruiter.com/recruiter-guide/hiring-glossary-a-to-z/job-displacement/>

-   replacement rate -
    <https://www.investopedia.com/terms/r/replacement-rate.asp>

-   slack work - <https://www.insee.fr/en/metadonnees/definition/c1974>

-   TakeUp rate -
    <https://financial-dictionary.thefreedictionary.com/Take+Up+Rate>

\#\#Reference Materials -
<https://oui.doleta.gov/unemploy/uifactsheet.asp>

Very good read -
<https://www.epi.org/publication/section-5-benefit-levels-increase-ui-benefits-to-levels-working-families-can-survive-on/>

Reference Paper -
<https://sci-hub.se/https://doi.org/10.1080/07350015.1995.10524593> \#\#
what makes people eligible for the Unemployment insurance benefits?

## Does any of the above terms relate??

Draw references from some papers to justify the analysis Paper 1 -
<https://www.jstor.org/stable/1392373> Paper 2 -

\#\#Load required packages

```{r pkgs, include=FALSE}
library(tidyverse)
```

## Data Importation and Cleaning

```{r import, echo=TRUE}
benefits <- read.csv("Benefits.csv", header = T)
benefits %>% head(10)
benefits %>% summary
benefits %>% dim
benefits <-  benefits[ ,-1] #redundant column
benefits %>% str
benefits %>% is.na %>% sum # no missing values
## On studying the data column 'bluecol', it seems to be redundant as all the workers are 'blue collar-ed'
data2 <- data2 %>% select(-(bluecol))


#Check the sixe of the dataset
object.size(data)
print(object.size(data), units = "Mb")
```

Several of the columns are in the character data type and would be
changed to factor. From the understanding of what each of the columns
represents we would be converting each column to the most appropriate
data type. e.g.

## Data Manipulation

-   Change "yrdispl" back to years and make a factor.
-   Change "state" to a factor data type.

```{r manipulate, echo=TRUE, warning=FALSE, paged.print=TRUE}
#Change specific columns by their name to factor
benefits2 <- benefits %>% mutate(across(c(state, joblost, nwhite, school12, sex,
                                          bluecol, smsa, married, dkids, dykids, 
                                          yrdispl, head, ui), factor))
str(benefits2)
```

\#\#Data Visualization Some visuals to explore the factors determining
the chances of an applicants' from receiving the Unemployment Insurance
benefits.

```{r datviz, echo=TRUE, message=FALSE, warning=FALSE}
benefits2 %>% ggplot(mapping = aes(x = joblost, fill = ui)) + 
  geom_bar(stat = "count") + ggtitle("UI applicants' status based on 'joblost' category")
```

```{r vizsex, echo=TRUE}
benefits2 %>% ggplot(mapping = aes(x = sex, fill = ui)) + 
  geom_bar(stat = "count") + ggtitle("UI applicants' status based on Gender")
```

```{r viz_married}
benefits2 %>% ggplot(mapping = aes(x = married, fill = ui)) + 
  geom_bar(stat = "count") + ggtitle("UI applicants' status based on marital status")
```

```{r viz_yrdispl}
benefits2 %>% ggplot(mapping = aes(x = yrdispl, fill = ui)) + 
  geom_bar(stat = "count") + ggtitle("UI applicants' status based on year of displacement")
```

## Data Modelling

###Simple Linear Regression
```{r}
sapply(mpg, levels)
str(mpg)
str(benefits2)
```

### Extensive Model assessments

### Predicting new values

### Perform assumptions check

## Build a multiple regression model

## Conclusion
